Classical Planning Algorithms ¡ on the Atari Video Games: Preliminary Results

### 2014-10-31 workshop-paper/experimental_results.tex
Third pass on the experimental results

We have tested the $\IW(1)$ and \BFS~described in the previous Section
over $9$ games selected amongst those considered in the \emph{training} set
of \cite{bellemare:jair2013}, and those discussed in \cite{deep-mind-atari}. 
In Table~\ref{performance} the performance of both $\IW(1)$ 
and \BFS, characterized by the average score obtained over all episodes, the standard deviation
of the score obtained and time (in seconds) per lookahead, is compared with that of \UCT, 
\BRFS~and the baseline random action selection. Our experimental 
setup  follows that of \cite{bellemare:jair2013}, with two exceptions. First, rather than
considering $10$ episodes of each game, we have considered just $5$
due to time constraints, as each game can take as much as $40$ hours of 
CPU time. Second, while \UCT~is limited to perform up to $500$ base policy rollouts ($150,000$ frames
are to be simulated while performing the lookahead), both $\IW(1)$ and \BFS are limited to
simulate $100,000$ frames. \UCT~parameters such as the exploration constant $C$ and
the planning horizon $m$ have not been modified ($C=0.1$, $m=300$), the latter being also
used by both $\IW(1)$ and \BFS~to limit the depth of the lookahead. We include
the average score reported for \UCT~by \cite{bellemare:jair2013} in Table~\ref{performance}
bracketed with parenthesis for reference. Also unchanged is the discount factor
used by all the algorithms ($\gamma=0.995$) and the rate at which an algorithm is
invoked to select an action (every $5$ frames). All algorithms reuse the tree built between each 
lookahead by setting the root of the tree to the best child of the root in the previous one,
and deleting its siblings and their descendants, as \cite{bellemare:jair2013} does for \UCT. We note that 
this may have a different effect on algorithms where all branches have the same number of
generated nodes, as is the case of \BRFS, than in $\IW(1)$, \BFS~and \UCT, where the depth of
the branches is uneven.

@@

We observe that in $4$ out of $9$ games, $\IW(1)$ outperforms \UCT~limited to the same number
of simulated frames, both in terms of average score attained and average time per call. \BFS~is
generally faster than $\IW(1)$, but only outperforms \UCT~in $2$ out of $9$ games (\Freeway and
\Seaquest) and $\IW(1)$ in $4$ out of $9$ games. Compared with the \UCT~results reported
in \cite{bellemare:jair2013}, that by limiting the number of rollouts allow a $50$\% more frames
to be simulated in each lookahead, $\IW(1)$ still outperforms \UCT~in $3$ out of $9$ games
(\Breakout, \Enduro and \Freeway). As introduced in the previous Section, the order in which
$\IW(1)$ and \BFS~consider actions is randomized, since the novelty of a given state may
change depending on the order in which states are considered. We found this randomization to
have a significant effect in \Qbert where not randomizing the order in which actions are considered
resulted in $\IW(1)$ being trapped in a loop. Last, we note that both $\IW(1)$ and \BFS~attain similar 
performance on \Freeway as the baseline semi-random algorithm \emph{Perturb} 
from \cite{bellemare:jair2013}, which beated both \UCT~and the reinforcement learning algorithms reported.

@@

We do not report $IW(2)$ in Table~\ref{performance} since we found its performance to
be very similar to that of \BRFS~in some preliminary tests. When considering $2$--tuples
of state variables to evaluate the novelty of states, it tended to result in lookaheads
$5$ times deeper than those resulting from applying \BRFS~, but still far too shallow so
as to compete with the other planning algorithms considered.

>>>>

We have tested the $\IW(1)$ and \BFS~described in the previous Section
over $9$ games selected amongst those considered in the \emph{training} set
of \cite{bellemare:jair2013}, and those discussed in \cite{deep-mind-atari}. 
In Table~\ref{performance} the performance of both $\IW(1)$ 
and \BFS, characterized by the average score obtained over all episodes, the standard deviation
of the score obtained and time (in seconds) per lookahead, is compared with that of \UCT, 
\BRFS~and the baseline random action selection. Our experimental 
setup  follows that of \cite{bellemare:jair2013}, with three exceptions. First, rather than
considering $10$ episodes of each game, we have considered just $5$
due to time constraints, as each game can take as much as $40$ hours of 
CPU time. Second, while \UCT~is limited to perform up to $500$ base policy rollouts ($150,000$ frames
are to be simulated while performing the lookahead), both $\IW(1)$ and \BFS are limited to
simulate $100,000$ frames. Other \UCT~parameters such as the exploration constant $C$ and
the planning horizon $m$ have not been modified ($C=0.1$, $m=300$). Thirdly and last, both $\IW(1)$ 
and \BFS~are limited to a planning horizon $m=1500$, where one only an action is evaluated every $5$ frames, matching
the setup used during actual gameplay (an action is selected every $5$ frames, using a no-op
for the remaining $4$). We include the average score reported for \UCT~by \cite{bellemare:jair2013} 
in Table~\ref{performance} bracketed with parenthesis for reference. Also unchanged is the discount factor
used by all the algorithms ($\gamma=0.995$). All algorithms reuse the tree built between each 
lookahead by setting the root of the tree to the best child of the root in the previous one,
and deleting its siblings and their descendants, as \cite{bellemare:jair2013} does for \UCT. We note that 
this may have a different effect on algorithms where all branches have the same number of
generated nodes, as is the case of \BRFS, than in $\IW(1)$, \BFS~and \UCT, where the depth of
the branches is uneven.

@@

We observe that in $4$ out of $9$ games, $\IW(1)$ outperforms \UCT~limited to the same number
of simulated frames, both in terms of average score attained and average time per call. \BFS~is
generally faster than $\IW(1)$, but only outperforms \UCT~in $2$ out of $9$ games (\Freeway and
\Seaquest) and $\IW(1)$ in $4$ out of $9$ games. Compared with the \UCT~results reported
in \cite{bellemare:jair2013}, that by limiting the number of rollouts allow a $50$\% more frames
to be simulated in each lookahead, $\IW(1)$ still outperforms \UCT~in $3$ out of $9$ games
(\Breakout, \Enduro~and \Freeway). As introduced in the previous Section, the order in which
$\IW(1)$ and \BFS~consider actions is randomized, since the novelty of a given state may
change depending on the order in which states are considered. We found this randomization to
have a significant effect in \Qbert~where not randomizing the order in which actions are considered
resulted in $\IW(1)$ being trapped in a loop. Last, we note that both $\IW(1)$ and \BFS~attain similar 
performance on \Freeway~as the baseline semi-random algorithm \emph{Perturb} 
from \cite{bellemare:jair2013}, which beated both \UCT~and the reinforcement learning algorithms reported.

@@

We do not report $IW(2)$ in Table~\ref{performance} since we found its performance to
be very similar to that of \BRFS~in some preliminary tests. When considering $2$--tuples
of state variables to evaluate the novelty of states, it tended to result in a lookahead
$5$ times deeper than those resulting from applying \BRFS~, but still far too shallow so
as to compete with the other planning algorithms considered.
